Viscous Erosion in Porous Media

### 2018-03-16 docs/methods/erosion_methods.tex
NM

To better understand the shape evolution, Fig.~\ref{shrink_intface}(a) shows several superimposed interfaces at evenly spaced times. The interfaces are color-coded by normalized time $t/t_f$, where $t_f = 1.79 \times 10^{-2}$ is the time at which the body vanishes in the simulation. The body converges to a fairly slender, for-aft symmetric shape with corners at its front and back --- somewhat reminiscent of an American football.
Fig.~\ref{shrink_intface}(b) shows how the absolute shear stress, $\atau$, is distributed along each of these interfaces. The normalized arclength $s/L$ runs counterclockwise around the body, with $s/L = $ 0 and 0.5 corresponding to the rear and front stagnation points respectively. At early times (yellow curves), the stress varies significantly, with more stress concentrated at the top and bottom of the body. As erosion alters the body's shape, though, the stress becomes more uniformly distributed over the entire surface (red curves). As the body nears the final stages of erosion (dark red), $\atau$ shows slight peaks near the stagnation points. Upon close inspection, the stress always vanishes {\em at} the stagnation points, but in a region that becomes increasingly narrow.

@@

The drag exerted by the body offers

The drag exerted by the eroding body

The drag on the eroding body offers a second check on our simulation. This quantity will also be of great interest in its own right for characterizing a porous-medium.

%\ref{eqn:drag}

In Fig.~\ref{area_drag}(b), we show the horizontal drag $F_D$ measured on the body as it erodes. Note that, in the context of 3D slender-body theory, this $F_D$ should be interpreted as the drag per unit span. The figure shows the drag to decrease  gradually at first, then abruptly tend to zero in the final moments as the body vanishes. This behavior is consistent with fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. We also show in the figure the separate contributions of the viscous and the pressure drag. Initially, these two contributions are comparable, with a slightly greater level of pressure drag. However, as the body erodes and assumes a more slender form, the pressure drag decays more rapidly and viscous drag dominates.

As a final check, we can estimate the drag (per unit span) using slender-body theory via the formula
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}.
\end{align}
The normalization factor of $4 \pi$ produces an {\em exact} relationship (under the assumptions of slender-body theory) for the case of a {\em circular} cylinder, where $a$ is taken to be the radius. To generalize to other shapes, we define $a = \sqrt{A/\pi}$ and fit $\ell$ to get $\ell = 2.3 a_0$.

\todo[inline]{Clean up this section}

\subsection{Local analysis for the limiting shape}

>>>>

To better understand the shape evolution, Fig.~\ref{shrink_intface}(a) shows several superimposed interfaces at evenly spaced times. The interfaces are color-coded by normalized time $t/t_f$, where $t_f = 1.79 \times 10^{-2}$ is the time at which the body vanishes in the simulation. As seen in Fig.~\ref{shrink_intface}(a), the body converges to a fairly slender, for-aft symmetric shape with corners at its front and back --- somewhat reminiscent of an American football.
Fig.~\ref{shrink_intface}(b) shows how the absolute shear stress, $\atau$, is distributed along each of these interfaces. The normalized arclength $s/L$ runs counterclockwise around the body, with $s/L = $ 0 and 0.5 corresponding to the rear and front stagnation points respectively. At early times (yellow curves), the stress varies significantly, with more stress concentrated at the top and bottom of the body. However, as erosion alters the body's shape, the stress becomes more uniformly distributed over the entire surface (red curves). The absolute stress $\atau$ shows slight peaks near the stagnation points as the bodies nears the final stages of erosion (dark red). Closer inspection reveals that the stress actually vanishes {\em at} the stagnation points, but in a region that becomes increasingly narrow with time.

@@

	The drag exerted by the eroding body offers a second check on the simulation. Figure~\ref{area_drag}(b) shows the horizontal drag, $F_D$, calculated in the simulations (black curve) using Eq.~\eqref{eqn:drag}. In the context of slender-body theory, $F_D$ should be interpreted as drag-per-unit-span. As seen in Fig.~\ref{area_drag}(b), the drag decreases gradually at first and then abruptly tends to zero in the very final moments. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. However,  as the body erodes and assumes a more slender form, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For the case of a {\em circular} cylinder, the normalization factor of $4 \pi$ produces an {\em exact} relationship (under the slender-body assumptions), where $a$ is taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale (since it generalizes the case of a circle). In Fig.~\ref{area_drag}(b) we show this scaling prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ the radius of the initially circular body. The scaling law agrees closely with the simulated drag as the body vanishes, thus providing validation.

\subsection{Local analysis for the limiting shape}
\label{LimitingShape}

\todo[inline]{Here in revising}
